---
title: "p8105_hw2_cl3938"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

Define relative path to the data
```{r path}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read in the Mr. Trashwheel dataset.

```{r trashwheel}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

```{r precipitation}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)

precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)

```

Combine the two precipitation datasets: 
```{r combine}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland.Mr. Trash Wheel collects litter and debris carried by the Jones Falls River toward the harbor, and to date has removed over a million pounds of litter. The dataset contains information on year, month, and volume trash collected, broken down by major trash types (for example, plastic bottle vs polystyrene amounts, number of chip bags).

There are a total of `r nrow(trashwheel_df)` rows in our final Mr. Trash Wheel dataset.

* An observation of interest: the median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`

Additional information regarding precipitation is also available in this dataset, primarily inches of precipitation broken down into year and month for the years 2017 & 2018.

There are a total of `r nrow(precip_df)` rows in our final Mr. Trash Wheel dataset.

* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.